### 2014-06-12 ijlyttle/first_pass.Rmd
ijl first set of plots

```{r load_libraries, message=FALSE}
library(stringr)
library(ggplot2)
library(dplyr)
library(reshape2)
library(lubridate)
library(ggvis)
```

@@

```{r twiddle}
df_twiddle <- student2012 %>%
  select(country = CNT, age = ST06Q01,
         math = PV1MATH, reading = PV1READ, science = PV1SCIE) %>%
  melt(id.vars = c("country", "age"), 
       variable.name = "subject", 
       value.name = "score")  %>%
  do(filter(., complete.cases(.))) %>%
  dplyr::group_by(country, subject) %>%
  summarize(n = n(), gini_coeff = ineq(score, type="Gini")) %>%
  arrange(as.character(country))
```

>>>>

```{r load_libraries, message=FALSE}
library(stringr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggvis)
```

@@

```{r twiddle}
df_twiddle <- student2012 %>%
  select(country = CNT, age = ST06Q01,
         math = PV1MATH, reading = PV1READ, science = PV1SCIE) %>%
  gather(key = subject, value = score, -country, -age) %>%
  do(filter(., complete.cases(.))) %>%
  dplyr::group_by(country, subject) %>%
  summarize(
    n = n(), 
    gini_coeff = ineq(score, type="Gini"), 
    mean = mean(score), 
    median = median(score)) %>%
  arrange(as.character(country))
```

"Alabama first" works to look for a specific country, but lets arrange by something more meaningful.

Let's get a mean of coefficents over all subjects, to help us order things.

```{r by_country}
df_twiddle_mean <- df_twiddle %>%
  group_by(country) %>%
  summarize(
    gini_coeff = mean(gini_coeff),
    mean = mean(mean),
    median = mean(median)
  )
```

```{r arrange}
my_levels <- 
  df_twiddle_mean$country[rev(order(df_twiddle_mean$gini_coeff))] %>%
  as.character

df_twiddle$country <- factor(
  df_twiddle$country,
  levels = my_levels
)

head(my_levels)

tail(my_levels)
```

I know I should figure out the `ggmap` package (I'll look at Di's code), but in the meantinme:

```{r vis_ineq_median}
ggplot(
  df_twiddle %>% filter(subject != "mean_subject"), 
  aes(x = gini_coeff, y = median, color = subject, size = n)
) + 
geom_point(alpha = 0.75) + 
scale_size_area()
```

This is print, so using ggvis and toolips is out of the question...

There appear two clusters - it will be interesting to see if there is a thread to bind them.

```{r vis_countries}
ggplot(
  df_twiddle,
  aes(x = gini_coeff, y = country, color = subject, size = n)
) + 
geom_point() + 
scale_size_area()

ggplot(
  df_twiddle,
  aes(x = median, y = country, color = subject, size = n)
) + 
geom_point() + 
scale_size_area()
```

Hopefully enough to get started.
